---
title: "Analysis"
format:
  html:
    embed-resources: true
editor: visual
execute: 
  echo: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_light())
library(ggdist)
library(readxl)
library(brms)
```

## Read data

```{r}
#| label: read-data

relatives_ae <- read_xlsx("data/Relatives.xlsx") %>%
  # Remove observations to be excluded
  filter(include == "y") %>%
  mutate(
    manuscript = "Aelfric"
  )

relatives_gdh <- read_xlsx("data/Relatives-2.xlsx", sheet = 3) %>%
  # Remove observations to be excluded
  filter(include == "y") %>%
  mutate(
    manuscript = "GDH"
  )

colnames(relatives_ae) <- c("text", "file", "parallel_set", "parallel_set_id", "dem", "mention", "include", "position", "correlative", "the", "divinity", "generic", "be", "manuscript")

colnames(relatives_gdh) <- c("text", "file", "parallel_set", "parallel_set_id", "dem", "mention", "include", "position", "correlative", "the", "divinity", "generic", "be", "antecedent", "manuscript")

relatives <- bind_rows(relatives_ae, relatives_gdh) %>%
  add_count(dem, divinity, name = "grand_tot") %>%
  mutate(
    dem = recode_factor(dem, n = "no", se = "SE", that = "THAT"),
    mention = recode_factor(mention, n = "other", y = "second"),
    position = recode_factor(position, n = "other", y = "early"),
    divinity = recode_factor(divinity, n = "other", y = "divinity"),
    mention_position = interaction(mention, position, lex.order = TRUE)
  )

relatives_counts <- relatives %>%
  filter(divinity == "other") %>%
  count(manuscript, dem, mention, position, grand_tot) %>%
  drop_na()
```

Check number of rows and columns

```{r}
#| label: checks

# Number of rows and columns
dim(relatives)
```

Glimpse of the data

```{r}
#| label: relatives
relatives
```

## Plotting

```{r}
#| label: counts

relatives %>%
  drop_na(mention_position, dem) %>%
  ggplot(aes(dem, fill = mention_position)) +
  geom_bar() +
  facet_grid(manuscript ~ divinity) +
  scale_fill_brewer(palette = "PRGn") +
  labs(
    title = "Number of relatives by mention and early/other verb"
  )
```

```{r}
#| label: proportions

relatives %>%
  drop_na(mention_position, dem) %>%
  ggplot(aes(dem, fill = mention_position)) +
  geom_bar(position = "fill") +
  facet_grid(manuscript ~ divinity) +
  scale_fill_brewer(palette = "PRGn") +
  labs(
    title = "Proportion of relatives by mention and early/other verb",
    y = "Proportion"
  )
```

## Modelling

```{r}
#| label: rel-bm

rel_bm <- brm(
  n ~
    dem * mention * position * manuscript +
    offset(log(grand_tot/100)),
  data = relatives_counts,
  family = poisson,
  backend = "cmdstanr",
  cores = 4,
  threads = threading(4),
  file = "data/rel_bm",
  prior = c(
    prior(normal(4, 1.5), class = Intercept),
    prior(normal(0, 2), class = b)
  ),
  seed = 9899
)
```

I fitted a Bayesian linear model using a Poisson distribution for the outcome variable (number of relatives), with demonstrative, second mention and early verb as predictors (including interactions thereof).

The following reports a summary of the estimated effects (in log-odds).

```{r}
#| label: rel-bm-effs

fixef(rel_bm) %>% as_tibble(rownames = "Term")
```

The following plot shows the predicted counts (based on a baseline of 100 relative phrases) as estimated by the model. The predicted counts can be also thought of as percentages (due to the 100-phrases baseline).

For each condition/group, the plot shows the predicted median count as a circle plus the 95% Credible Intervals (CrI) as an error bar. The 95% CrIs can be interpreted as: "There is a 95% probability that the true count value is within this interval, based on the data and model". Note that inference should be based on the CrI rather than just on the median value.

```{r}
#| label: rel-bm-conds

conds <- make_conditions(relatives_counts, c("manuscript", "dem")) %>%
  mutate(grand_tot = 100)

conditional_effects(rel_bm, effects = "position:mention", conditions = conds)
```

```{r rel-bm-cond}
rel_bm_draws <- as_draws_df(rel_bm)

new_data <- expand_grid(
  dem = c("no", "SE", "THAT"),
  mention = c("other", "second"),
  position = c("other", "early"),
  manuscript = c("Aelfric", "GDH"),
  grand_tot = 100
)

headings <- new_data %>%
  unite("heads", dem:manuscript, sep = ".") %>%
  pull(heads)

rel_bm_cond <- posterior_epred(rel_bm, newdata = new_data)
colnames(rel_bm_cond) <- headings

rel_bm_cond <- as_tibble(rel_bm_cond)

rel_bm_cond_l <- rel_bm_cond %>%
  pivot_longer(everything(), names_to = "condition") %>%
  separate(
    condition,
    c("dem", "mention", "position", "manuscript"),
    sep = "\\."
  ) %>%
  arrange(manuscript, dem, mention, position)
```

```{r rel-bm-cond-ael}
rel_bm_cond_l %>%
  filter(manuscript == "Aelfric") %>%
  ggplot(aes(dem, value)) +
  stat_interval(.width = c(0.5, 0.7, 0.9, 0.99), linewidth = 5) +
  facet_grid(position ~ mention, labeller = label_both) +
  coord_cartesian(ylim = c(0, 50)) +
  scale_color_brewer(direction = -1) +
  labs(
    title = "Aelfric manuscript",
    x = "Demonstrative", y = "Percentage"
  )
```

```{r rel-bm-cond-gdh}
rel_bm_cond_l %>%
  filter(manuscript == "GDH") %>%
  ggplot(aes(dem, value)) +
  stat_interval(.width = c(0.5, 0.7, 0.9, 0.99), linewidth = 5) +
  facet_grid(position ~ mention, labeller = label_both) +
  coord_cartesian(ylim = c(0, 50)) +
  scale_color_brewer(direction = -1) +
  labs(
    title = "GDH manuscript",
    x = "Demonstrative", y = "Percentage"
  )
```
